The Sub-Millimetre User Reduction Facility: an iterative map-maker for SCUBA-2

### 2012-01-11 smurf_paper.tex
add important map-making ref + downsampling section draft

There are numerous ways to attack this problem, most of them evolving
out of techniques that were developed to produce maps of the
large-scale anisitropies in the Cosmic Microwave Background (CMB).  Styles
of map-making: maximum likelihood \citep[e.g.,][]{patanchon2008};
iterative -- see \citet{johnstone2000} implementation of
\citet{wright1996} to remove chop, \citet{kovacs2008} for SHARC-2
etc.; de-correlation using PCA analysis
\citep[e.g.][]{laurent2005,scott2008,aguirre2010}.

@@

\subsubsection{Down-sampling}
\label{sec:downsamp}

>>>>

There are numerous ways to attack this problem, most of them evolving
out of techniques that were developed to produce maps of the
large-scale anisitropies in the Cosmic Microwave Background (CMB).
Styles of map-making: least squares \citep{janssen1992} maximum
likelihood \citep[e.g.,][]{patanchon2008}; iterative -- see
\citet{johnstone2000} implementation of \citet{wright1996} to remove
chop, \citet{kovacs2008} for SHARC-2 etc.; de-correlation using PCA
analysis \citep[e.g.][]{laurent2005,scott2008,aguirre2010}.

@@

\subsubsection{Time-series down-sampling and map pixel size}
\label{sec:downsamp}

The highest useful frequency in the nominally 200\,Hz-sampled
\scuba\ data is that which corresponds to the smallest angular scale
that the instrument is sensitive to. As mentioned earlier, the usual
rule-of-thumb for a Gaussian beam is to provide at least 3 samples for
each FWHM, or roughly 2.5\,arcsec for the 7.5\,arcsec
450\,\micron\ channel, and 5\,arcsec for the 14.5\,arcsec
850\,\micron\ channel. For a typical scan speed of
300\,arcsec\,s$^{-1}$, the maximum useful sample rate is therefore
about 120\,Hz at 450\,\micron, and 40\,Hz at 850\,\micron. In order to
save execution time and memory usage (both of which scale linearly
with data volume), it is clearly advantageous to re-sample the data to
these lower rates.

Similarly, from the point of view of retaining all of the useful
information, the size of the output map pixels need not be any smaller
than the FWHM/3 rule. Larger pixels may be selected, but at the
expense of retaining information at the diffraction limit of the
telescope.

The default behaviour of SMURF is to select output map pixel sizes of
2 and 4\,arcsec at 450 and 850\,\micron, respectively (slightly
over-sampled). The slew speed of the telescope (approximately constant
for all standard scan modes) is then used to calculate the highest
useful frequency for the time-series given the map pixel size,
$f_\mathrm{max}$. The data are then re-sampled to this new rate prior
to map-making (as well as the data cleaning steps).

The method used by SMURF to down-sample simply averages together
multiple samples from the original time-series, $x_i$ to estimate the
lower-frequency output time-series, $y_j$. In general there will not
be an integer number of samples from $x_i$ in each of the $y_j$, so
fractional samples from $x_i$ are used at the $y_j$ time
boundaries. The algorithm is fast since each sample in $x_i$ need only
be visited once. From a spectral point of view, this boxcar average is
equivalent to applying a sinc-function low-pass filter in frequency
space. Such behaviour is desirable since it serves as an anti-aliasing
filter; most non-white features above the Nyquist frequency in $y_j$
that could contaminate the output data are largely removed.

As an alternative, we also investigated an algorithm in which the FFT
of $x_i$ is simply truncated to the target sample rate before
transforming back to the time domain (i.e., applying a hard-edged
low-pass filter). In practice the noise performance was
indistinguishable, and required slightly more execution time, using
this method compared to the boxcar method described above. It is
therefore not currently available to users.
